---
title: "ResIN_ANES_timeseries_1"
author: "Us versus them"
date: "2023-07-06"
output: html_document
---

```{r import, include=FALSE, warning = false, message=FALSE}

options(collapse_mask = "manip")
rm(list = setdiff(ls(), lsf.str()))

if(!require("pacman")) install.packages('pacman')
library(pacman)
p_load(ResIN)
p_load(tidyverse)
p_load(patchwork)
p_load(qgraph)
p_load(ggpubr)

anes <- read.csv("anes_out.csv")

node_covars <- c("lib_con", "pid", "egal_1", "egal_2", "auth_1", "auth_2", "auth_3", "auth_4", "rac_res_1", "rac_res_2", "rac_res_3", "rac_res_4", "ft_jews", "ft_blacks", "ft_whites", "ft_bis", "ft_lab" , "ft_lib" , "ft_con" , "ft_his" , "ft_dem" , "ft_poor" , "ft_rep" , "ft_asian" , "ft_congr" , "ft_gay" , "ft_immigr", "ft_chrfund" , "ft_fem" , "ft_mus" , "ft_chr" , "hw_whites" , "hw_blacks", "hw_his", "hw_asian")

node_costats <- rep("mean", length(node_covars))

anes_sub <- anes %>% dplyr::select("year", "gov_health", "guar_jobs", "abort", "def_spend", "aid_black", "spend_serv", node_covars)
```

```{r}

anes_sub$aff_pola <- anes_sub$ft_dem - anes_sub$ft_rep


anes_2020 <- anes_sub %>%  filter(year==2020) %>% select(!year)


ResIN_out_strength <- anes_2020 %>% ResIN(node_vars = c("gov_health", "guar_jobs", "abort", "aid_black", "spend_serv"),
                                          left_anchor = "abort_4", plot_whichstat = "Strength",  plot_responselabels = T, plot_ggplot = F,
                                          color_palette = "Greens", plot_title = "ResIN: Strength centrality", seed = 22)


ResIN_out_closeness <- anes_2020 %>% ResIN(node_vars = c("gov_health", "guar_jobs", "abort", "aid_black", "spend_serv"),
                                          left_anchor = "abort_4", plot_whichstat = "Closeness",  plot_responselabels = T, plot_ggplot = F,
                                          color_palette = "Reds", plot_title = "ResIN: Closeness centrality", seed = 22)


ResIN_out_betweenness <- anes_2020 %>% ResIN(node_vars = c("gov_health", "guar_jobs", "abort", "aid_black", "spend_serv"),
                                          left_anchor = "abort_4", plot_whichstat = "Betweenness",  plot_responselabels = T, plot_ggplot = F,
                                          color_palette = "Blues", plot_title = "ResIN: Betweenness centrality", seed = 22)

```

```{r, warning=FALSE, message=FALSE}
## Making BNA - graph
bna_out <- anes_2020 %>%  select(c("gov_health", "guar_jobs", "abort", "aid_black", "spend_serv")) %>% cor_auto() %>% abs() %>% as.data.frame()
diag(bna_out) <- 0

bna_igraph <- igraph::graph_from_adjacency_matrix(as.matrix(bna_out), mode = "undirected", weighted = TRUE, diag = FALSE)
bna_layout <- as.data.frame(prcomp(igraph::layout_nicely(bna_igraph))$x)

bna_layout$node_names <- colnames(bna_out)
colnames(bna_layout) <- c("x", "y", "node_names")

bna_layout$y <- bna_layout$y*3
bna_layout$x <- bna_layout$x*3.5

node_frame <- bna_layout
node_frame$from <- node_frame$node_names

bna_qgraph <- qgraph::qgraph(bna_out, DoNotPlot = TRUE, layout = "spring", labels = rownames(bna_out))

node_net_stats <- qgraph::centrality_auto(bna_qgraph, weighted = TRUE)
node_frame <- cbind(node_frame, node_net_stats$node.centrality)

g <- igraph::as_data_frame(bna_igraph)
g$from.x <- node_frame$x[match(g$from, node_frame$node_names)]
g$from.y <- node_frame$y[match(g$from, node_frame$node_names)]
g$to.x <- node_frame$x[match(g$to, node_frame$node_names)]
g$to.y <- node_frame$y[match(g$to, node_frame$node_names)]

edgelist_frame <- dplyr::left_join(g, node_frame, by = "from")

BNA_ggplot <- ggplot2::ggplot()+ ggplot2::coord_fixed(ratio=1, x = c(min(edgelist_frame$x-1.5), max(edgelist_frame$x+1.5)),
                y = c(min(edgelist_frame$y-1.5), max(edgelist_frame$y+1.5)))+
  ggtitle("BNA: Node centrality (strength & closeness)")

BNA_ggplot <- BNA_ggplot + ggplot2::geom_curve(ggplot2::aes(x = edgelist_frame$from.x, xend = edgelist_frame$to.x, y = edgelist_frame$from.y, yend = edgelist_frame$to.y, linewidth = edgelist_frame$weight), curvature = 0.2, color = "black", alpha = 0.4) +
  ggplot2::scale_linewidth(range = c(0, 2))

BNA_ggplot <- BNA_ggplot + ggplot2::geom_text(ggplot2::aes(x = node_frame$x, y = node_frame$y, label = node_frame$node_names, alpha = node_frame$Strength), size = 4) +
  scale_alpha_continuous(range = c(0.6, 1.5))

BNA_ggplot <- BNA_ggplot+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.line = ggplot2::element_blank(), axis.text.x = ggplot2::element_blank(), axis.title.x = ggplot2::element_blank(),
          axis.text.y = ggplot2::element_blank(), axis.title.y = ggplot2::element_blank(),
          axis.ticks = ggplot2::element_blank(), panel.grid.major = ggplot2::element_blank(),
          panel.grid.minor = ggplot2::element_blank(), legend.position = "none",
          legend.text = ggplot2::element_blank(), plot.title = ggplot2::element_text(hjust = 0.5, size=12, vjust = -7.5), plot.margin = margin(-2,0,0,0))

str_plot <- ResIN_out_strength$ResIN_ggplot + theme(legend.position = "none", plot.title = element_text(size = 12, vjust = -9), plot.margin = margin(-2,0,0,0))
cl_plot <- ResIN_out_closeness$ResIN_ggplot + theme(legend.position = "none", plot.title = element_text(size = 12, vjust = -9), plot.margin = margin(-2,0,0,0))
bet_plot <-  ResIN_out_betweenness$ResIN_ggplot + theme(legend.position = "none", plot.title = element_text(size = 12, vjust = -9), plot.margin = margin(-2,0,0,0))

combi <- ggarrange(BNA_ggplot, str_plot, cl_plot, bet_plot, ncol = 1, nrow = 4)

ggsave("combi.pdf",combi, device = "pdf", width = 8, height = 8)


```

A comparison of node strength centrality in ResIN can be indicative of the relative importance of different attitudes in either belief cluster. As seen in the top pane of Figure XX, defending abortion rights issue is much more central attitude to Liberals while the most important issues for conservatives deal with government sponsored healthcare and federal job guarantees. Such a result would appears surprising from the perspective of classic BNA, which would simply locate the guaranteed jobs item as the most central overall, remaining oblivious to the possibility that different attitudes can be more or less central to different ideological communities.

One quirky feature of classic BNA is that strength and closeness centrality are equivalent for networks in which all shortest network paths are direct paths. Furthermore, betweenness-centrality is constant remains constant at zero across all nodes. Figure XX shows that this is the case if we investigate our running example network with classic BNA model. In ResIN, however, closeness and betweenness centrality not only indicate different belief system functions, they are also largely decoupled from equivalencies with strength centrality.

Peeking at the middle pain in Figure XX, we note that nodes high closeness tend to lie in between the liberal and conservative clusters. These attitudes are typically moderate in nature and tend to be endorsed by people across the ideological aisle. However, not all beliefs that are well-connected to other beliefs actually help transmit information (or trust) across the most important communities within a belief network. In other words, only those attitudes which are also high in betweenness centrality likely act as gate-keepers or brokers between ideological sub-clusters. In a ResIN network, these nodes may be particularly relevant to de-polarization efforts. According to our model in the bottom pane of Figure XX, if you would like to find common ground between liberals and conservatives, your best bet would be to establish conversations about the general state of public services and about government healthcare. By contrast, you'd better avoid mentioning issues concerning abortion and race. The revealed patterns of co-endorsed beliefs suggest that liberals and conservatives may not agree on spending and health care any time soon, but their communities are (still) connected by sufficient moderate opinion holders that they might see reason and value in each others view-points.

```{r first_net, include=FALSE, warning = false, message=FALSE}

## 1988

net_1988 <- ResIN::ResIN(anes_sub[anes_sub$year==1988, ], node_vars = c("gov_health","guar_jobs", "abort", "def_spend", "aid_black","spend_serv"),  network_stats=TRUE, node_covars = node_covars, node_costats = node_costats, remove_negative = TRUE)
net_1988$node_frame$year <- 1988
net_1988$edgelist_frame$year <- 1988

if(net_1988$node_frame$x[net_1988$node_frame$node_names=="abort_1"] < 0 & net_1988$node_frame$x[net_1988$node_frame$node_names=="spend_serv_1"] < 0) {
      net_1988$edgelist_frame$from.x <- -1*(net_1988$edgelist_frame$from.x)
      net_1988$edgelist_frame$to.x <- -1*(net_1988$edgelist_frame$to.x)
      net_1988$edgelist_frame$x <- -1*(net_1988$edgelist_frame$x)
      net_1988$node_frame$x <- -1*(net_1988$node_frame$x)}

net_1992 <- ResIN::ResIN(anes_sub[anes_sub$year==1992, ], node_vars = c("gov_health","guar_jobs", "abort", "def_spend", "aid_black","spend_serv"),  network_stats=TRUE, node_covars = node_covars, node_costats = node_costats)
net_1992$node_frame$year <- 1992
net_1992$edgelist_frame$year <- 1992

if(net_1992$node_frame$x[net_1992$node_frame$node_names=="abort_1"] < 0 & net_1992$node_frame$x[net_1992$node_frame$node_names=="spend_serv_1"] < 0) {
      net_1992$edgelist_frame$from.x <- -1*(net_1992$edgelist_frame$from.x)
      net_1992$edgelist_frame$to.x <- -1*(net_1992$edgelist_frame$to.x)
      net_1992$edgelist_frame$x <- -1*(net_1992$edgelist_frame$x)
      net_1992$node_frame$x <- -1*(net_1992$node_frame$x)}

net_1996 <- ResIN::ResIN(anes_sub[anes_sub$year==1996, ], node_vars = c("gov_health","guar_jobs", "abort", "def_spend", "aid_black","spend_serv"),  network_stats=TRUE, node_covars = node_covars, node_costats = node_costats)

if(net_1996$node_frame$x[net_1996$node_frame$node_names=="abort_1"] < 0 & net_1996$node_frame$x[net_1996$node_frame$node_names=="spend_serv_1"] < 0) {
      net_1996$edgelist_frame$from.x <- -1*(net_1996$edgelist_frame$from.x)
      net_1996$edgelist_frame$to.x <- -1*(net_1996$edgelist_frame$to.x)
      net_1996$edgelist_frame$x <- -1*(net_1996$edgelist_frame$x)
      net_1996$node_frame$x <- -1*(net_1996$node_frame$x)}

net_1996$node_frame$year <- 1996
net_1996$edgelist_frame$year <- 1996

net_2000 <- ResIN::ResIN(anes_sub[anes_sub$year==2000, ], node_vars = c("gov_health","guar_jobs", "abort", "def_spend", "aid_black","spend_serv"),  network_stats=TRUE, node_covars = node_covars, node_costats = node_costats)

if(net_2000$node_frame$x[net_2000$node_frame$node_names=="abort_1"] < 0 & net_2000$node_frame$x[net_2000$node_frame$node_names=="spend_serv_1"] < 0) {
      net_2000$edgelist_frame$from.x <- -1*(net_2000$edgelist_frame$from.x)
      net_2000$edgelist_frame$to.x <- -1*(net_2000$edgelist_frame$to.x)
      net_2000$edgelist_frame$x <- -1*(net_2000$edgelist_frame$x)
      net_2000$node_frame$x <- -1*(net_2000$node_frame$x)}

net_2000$node_frame$year <- 2000
net_2000$edgelist_frame$year <- 2000

net_2004 <- ResIN::ResIN(anes_sub[anes_sub$year==2004, ], node_vars = c("gov_health","guar_jobs", "abort", "def_spend", "aid_black","spend_serv"),  network_stats=TRUE, node_covars = node_covars, node_costats = node_costats)

if(net_2004$node_frame$x[net_2004$node_frame$node_names=="abort_1"] < 0 & net_2004$node_frame$x[net_2004$node_frame$node_names=="spend_serv_1"] < 0) {
      net_2004$edgelist_frame$from.x <- -1*(net_2004$edgelist_frame$from.x)
      net_2004$edgelist_frame$to.x <- -1*(net_2004$edgelist_frame$to.x)
      net_2004$edgelist_frame$x <- -1*(net_2004$edgelist_frame$x)
      net_2004$node_frame$x <- -1*(net_2004$node_frame$x)}

net_2004$node_frame$year <- 2004
net_2004$edgelist_frame$year <- 2004

net_2008 <- ResIN::ResIN(anes_sub[anes_sub$year==2008, ], node_vars = c("gov_health","guar_jobs", "abort", "def_spend", "aid_black","spend_serv"),  network_stats=TRUE, node_covars = node_covars, node_costats = node_costats)

if(net_2008$node_frame$x[net_2008$node_frame$node_names=="abort_1"] < 0 & net_2008$node_frame$x[net_2008$node_frame$node_names=="spend_serv_1"] < 0) {
      net_2008$edgelist_frame$from.x <- -1*(net_2008$edgelist_frame$from.x)
      net_2008$edgelist_frame$to.x <- -1*(net_2008$edgelist_frame$to.x)
      net_2008$edgelist_frame$x <- -1*(net_2008$edgelist_frame$x)
      net_2008$node_frame$x <- -1*(net_2008$node_frame$x)}

net_2008$node_frame$year <- 2008
net_2008$edgelist_frame$year <- 2008

net_2016 <- ResIN::ResIN(anes_sub[anes_sub$year==2016, ], node_vars = c("gov_health","guar_jobs", "abort", "def_spend", "aid_black","spend_serv"),  network_stats=TRUE, node_covars = node_covars, node_costats = node_costats)

if(net_2016$node_frame$x[net_2016$node_frame$node_names=="abort_1"] < 0 & net_2016$node_frame$x[net_2016$node_frame$node_names=="spend_serv_1"] < 0) {
      net_2016$edgelist_frame$from.x <- -1*(net_2016$edgelist_frame$from.x)
      net_2016$edgelist_frame$to.x <- -1*(net_2016$edgelist_frame$to.x)
      net_2016$edgelist_frame$x <- -1*(net_2016$edgelist_frame$x)
      net_2016$node_frame$x <- -1*(net_2016$node_frame$x)}

net_2016$node_frame$year <- 2016
net_2016$edgelist_frame$year <- 2016

net_2020 <- ResIN::ResIN(anes_sub[anes_sub$year==2020, ], node_vars = c("gov_health","guar_jobs", "abort", "def_spend", "aid_black","spend_serv"),  network_stats=TRUE, node_covars = node_covars, node_costats = node_costats)

net_2020$node_frame$year <- 2020
net_2020$edgelist_frame$year <- 2020

if(net_2020$node_frame$x[net_2020$node_frame$node_names=="abort_1"] < 0 & net_2020$node_frame$x[net_2020$node_frame$node_names=="spend_serv_1"] < 0) {
      net_2020$edgelist_frame$from.x <- -1*(net_2020$edgelist_frame$from.x)
      net_2020$edgelist_frame$to.x <- -1*(net_2020$edgelist_frame$to.x)
      net_2020$edgelist_frame$x <- -1*(net_2020$edgelist_frame$x)
      net_2020$node_frame$x <- -1*(net_2020$node_frame$x)}

anes_meta_edge <- rbind(net_1988$edgelist_frame, net_1992$edgelist_frame, net_1996$edgelist_frame, net_2000$edgelist_frame, net_2004$edgelist_frame, net_2008$edgelist_frame, net_2016$edgelist_frame, net_2020$edgelist_frame)
anes_meta_node <- rbind(net_1988$node_frame, net_1992$node_frame, net_1996$node_frame, net_2000$node_frame, net_2004$node_frame, net_2008$node_frame, net_2016$node_frame, net_2020$node_frame)

write.csv(anes_meta_edge, "anes_meta_edge.csv")
write.csv(anes_meta_node, "anes_meta_node.csv")

```
